
I write this section after five years working in industry, coming back to university to finish my thesis gives me a feeling of coming home. It is impossible to come to this stage without the help of my main advisor, Prof. Bengt Jonsson, who is not only my PhD supervisor, but also my main advisor for my master's thesis, under his guidance, I have learned a lot from solving problems independently, writing papers as well as presentation skills.  I cannot thank him enough for his time working with me to complete this thesis.

My gratitude also goes to my second advisor, Prof. Parosh Aziz Abdulla, for his trust in me and his encouragement to try out new ideas, keep me motivated, and stay positive in any situation. I am deeply grateful for all his guidance and support.
My thanks also go to the previous and current members (in alphabetical order) of the Verification: Phi Diep Bui, Frédéric Haziza, Carl Leonardsson, Othmane Rezine, Amendra Shrestha, Jari Stenman, Tuan Phong Ngo and Yunyun Zhu. I will always remember the Friday lunches with you even though I always have problems sitting in Parosh's car to the restaurants.
I would also like to thank many people from the department who supported me during my study. In particular, thanks to
Magnus for helping me with the Swedish summarisation of this thesis. Thanks to Prof. Anders Berglund and Prof. Roland Bol for your help during my master's program study here in Uppsala.
This list would not be complete without mentioning my lifelong Vietnamese friends, we are a small Vietnam inside big Uppsala! Phong - Quỳnh Anh, Điệp - Dịu, Phúc - Trâm Anh, Thành - Diễm,
Lâm - Thùy, Minh - Thủy, Tuấn - Oanh, Long - Thái, Tuân - Liên, Hoàng
Long, Tuấn Anh, Nam, Lan, Thảo, Hoàng Tuấn, Thư, Dung, Mận, Viện - Nga, thank you for always being here for me.
Finally, I would like to express my deepest gratitude and love to my family. It's hard to forget and describe in words when parents had to cycle across communes to borrow money to prove their finances for their son who wasn't good enough to apply for a scholarship but wanted to study abroad, and there were many hardships. that my family had to go through during the time I studied here, to get to this point of writing this thesis, I can only say two words to thank my parents and sister very much.

The last words are for my small family, Ha, and Khi, the wife who always has a stomach ache every time her husband is right about career opportunities, but difficulties, just a little bit of hardship from him, is enough to understand her companionship. feel my wife. Thank you so much, wife!

